Type-Safe Client-Server Communication in Web Applications

### 2014-02-27 paper.tex
Remove redundant paragraph in intro.

Having your application rely on some network protocol for
its internal functionality involves writing a lot of boilerplate network code
and some serious advance planning of your application's features to be able to
design a suitable protocol, hampering incremental development. Were we to
discount this initial design burden, ensuring that the protocol is used
correctly throughout the code base as the application evolves and gains new
features would still put an ever increasing burden on the developer.

This state of things is not a concious design choice - most web applications
are conceptually a single entity, not two programs which just happen to talk
to each other over a network - but a consequence of there being a large,
distributed network between the client and server parts.
However, such implementation details should not be allowed to dictate the way
we structure and reason about our applications - clearly, an abstraction is
called for.

To put make all a bit more concrete, let's take a look at an example. Let's say
what we want to implement a simple ``chatbox'' component for a website, to allow
visitors to discuss the site's content in real time. Using mainstream
development practives and recent technologies such as WebSockets, we may come
up with something like the program in figure \ref{lst:javascript-client}.

@@

Messages are transmitted as strings, with the initial four characters
indicating the type of the message and the rest being the optional payload.
There are only two messages; a handshake indicating that a user wants to join
the conversation, and a broadcast message which sends a line of text to all
connected users via the server.

@@

Moreover, this code is not very extensible. If this simple application is to be
enhanced with new features down the road, the network protocol will clearly
need to be redesigned. However, if we were developing this application for a
client, said client would likely not want to pay the added cost for the design
and implementation of features she did not ask for, and perhaps never will.

@@

>>>>

This state of things is not a conscious design choice - most web applications
are conceptually a single entity, not two programs which just happen to talk
to each other over a network - but a consequence of there being a large,
distributed network between the client and server parts.
However, such implementation details should not be allowed to dictate the way
we structure and reason about our applications - clearly, an abstraction is
called for.

To put make the problem a bit more concrete, let's take a look at an example.
Let's say what we want to implement a simple ``chatbox'' component for a
website, to allow visitors to discuss the site's content in real time. Using
mainstream development practives and recent technologies such as WebSockets,
we may come up with something like the program in figure
\ref{lst:javascript-client}.

@@

Messages are transmitted as strings, with the initial four characters
indicating the type of the message and the rest being the optional payload.
There are only two messages; a handshake indicating that a user wants to join
the conversation, and a broadcast message which sends a line of text to all
connected users via the server. The only messages received from the server are
new chat messages, delivered as simple strings.

@@

Moreover, this code is not very extensible. If this simple application is to be
enhanced with new features down the road, the network protocol will clearly
need to be redesigned. However, if we were developing this application for a
client, said client would likely not want to pay the added cost for the design
and implementation of features she did not - and perhaps never will - ask for.

@@

Now that we have seen how to implement both network communication, we are ready
to revisit the chatbox program, this time using our improved programming model.
